---
title: "Final_Project: Who is King of Attacks?"
author: "Meaghan Barrett"
date: "2025-04-08"
output: 
    html_document:
        theme: paper
        highlight: tango
        toc: true
        toc_float:
            collapsed: true
        number_sections: false
        code_download: true
        df_print: kable
        code_folding: show
        mode: selfcontained
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
```

[Return to Homepage](../index.html)

# **Introduction**
In the following report, I will investigate and examine which animal is the king of attacks, by comparing the attacks from sharks, wolves, and alligators. While this data has already been cleaned and contains fairly simple information, this final project aims to investigate the disparities between these three animals, the attack locations, and the effected population. In simpler terms, who is responsible for the most attacks? Which populations are the most effected? Where these attacks provoked or not? These are just some of the questions I hope to explore in this writing. 

# **Background**
This project was inspired in part by my family, my mother is from Florida and there have been some very interesting documentaries done on all these animals. It was also inspired by general happen stance: I found a very nice data set on Kaggle and felt it could be an interesting take on a seemingly dull data set. As mentioned, this data set was found 
[link](https://www.kaggle.com/datasets/danela/fatal-alligator-attacks-us?resource=download&select=fatal_alligator_attacks_US.csv), as updated 2 years ago (2023),

[link](https://www.kaggle.com/datasets/felipeesc/shark-attack-dataset?phase=FinishSSORegistration&returnUrl=%2Fdatasets%2Ffelipeesc%2Fshark-attack-dataset%2Fversions%2F1%3Fresource%3Ddownload&SSORegistrationToken=CfDJ8PHSCL9k9s1HuJ2cRFBFhuhgpxN0g_ATDITz_-cXVG-n5-S8PcAnZdgDXHbn7ud0iaVYLeYWkYnFTY6Nc4JFt1nyWAZsTuhR8vSPv3ok5TP4AtRRK9-IzGDqSzZKUGxMayKK5NKkdWgewUVYPMF1aJl4phPB4ObwXl2AK7698CE230yss9kgbAVKcZACBg00FmSPPkTsYGhlWu4z3VrezvZDXoLn2eYayI0784JDAnaa1L5KVsvpzolGTk9T8hn7uDtX29rwNRaQWy19BsV0KZ7TcfDfFpYvRD8rSMrq4yEul7-CRa2L1R5qWvxEOYMlGI-VFN87sgabOPrg_CJ6jcJaVo0CcsQ&DisplayName=Meaghan+Barrett),as updated 4 years ago (2021),

[link](https://www.kaggle.com/code/icecream4/shark-attacks-by-hemispheres/notebook), as updated 8 years ago (2017),

[link](https://www.kaggle.com/datasets/danela/global-wolf-attacks?select=global_wolves.csv), as updated 2 years ago (2023)

The data set consists of various reported statistics regarding these different attacks. Most individuals were attacked were predatory. However, some individuals were reported to provoke the animal. Different age groups were attacked each time. 

Prevention would be great, but there is not signs for those that will and could be attacked. Alligators, sharks, and wolves have different reasons to attack.  

# **Data**
```{r, echo = FALSE, results = 'hide'}
# load all necessary libraries 
library(tidyverse)
library(janitor)
library(leaflet)
library(dplyr)
library(ggplot2)

# reading in kaggle data sets 
gator <- read.csv("fatal_alligator_attacks_US.csv")

g_wolves <- read.csv("global_wolves.csv")

shark_2 <- read.csv("shark_attacks.csv")

shark_1 <- read.csv("Shark_attacks/attacks.csv")

shark_3 <- read.csv("Shark_attacks/list_coor_australia.csv")

```
This data set was retrieved from Kaggle and has already been **somewhat** cleaned for analysis. However, there are some changes I wanted to make to the data structure, changing missing values/empty spaces, cleaning the names, and improving characters found throughout the data set. The following code and outputs demonstrate the changes I've made to allow for smoother data analysis: 

```{r, echo=TRUE}

# shows the original data structure 
str(gator)

str(shark_1)

str(shark_2)

str(shark_3)

# convert column titles to snake_case 

gator <-clean_names(gator)

g_wolves <-clean_names(g_wolves)

shark_2 <-clean_names(shark_2)

# converts all N/A chr values to actual missing values and fills in empty spaces with missing values 

g_wolves$type_of_attack[is.na(g_wolves$type_of_attack) | g_wolves$type_of_attack == ""] <- "Unknown"

shark_1$type[is.na(shark_1$type) | shark_1$type == "Invalid"] <- "Unknown"
shark_1_subset <- shark_1[1:6302, ]

shark_2$type[is.na(shark_2$type) | shark_2$type == "Invalid"] <- "Unknown"


# converts data structure to more appropriate data types

gator <-gator %>% 
  mutate(location = str_extract(details, "(Miami|Florida|Georgia|Texas|Louisiana|South Carolina)"),
         location = ifelse(location == "Miami", "Florida", location))

colnames(shark_3) <- c("latitude", "longitude")

```
```{r, echo = FALSE}
# Alligator

gator %>%
  mutate(date = as.Date(date, format = "%B %d, %Y")) %>%  
  filter(age != "?") %>%
  mutate(age = as.numeric(age)) %>%
  filter(age >= 2 & age <= 81) %>%
  mutate(year = as.integer(format(date, "%Y"))) %>%  
  arrange(age) %>%
  group_by(location) %>%
  ggplot(aes(x = factor(year), y = age, color = location)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Year", y = "Age")

deaths <-gator %>%
  filter(location != "?") %>%  
  filter(sex != "?") %>%       
  group_by(location, sex) %>%
  tally()

print(deaths)



deaths %>% 
  ggplot(aes(x = location, y = n, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +  
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Alligator Attacks by State",
       x = "State", y = "Number of Victims", fill = "Sex")

```
